customLayout
========================================================

```{r setup, include=FALSE}
# set global chunk options
library(knitr)
opts_chunk$set(cache=TRUE)
```

`customLayout` is a simple extension of basic `layout` function from `R`.

## Instalation:

`customLayout` is available only via GitHub and can be installed using `devtools`:

```{r, eval=FALSE}
require(devtools)
install_github("customLayout",username="zzawadz")
```


## Basic functionality:

You can simply create layouts like in base R:

```{r base, message = FALSE, results = 'hide', fig.width = 4, fig.height = 2, fig.keep = 'all', fig.pos = 'H'}
library(customLayout)
lay <- layCreate(
  matrix(1:4, nc = 2),
  widths = c(3, 2),
  heights = c(2, 1))
layShow(lay)

lay2 <- layCreate(
  matrix(1:4, nc = 2),
  widths = c(3, 5),
  heights = c(2, 4))
layShow(lay2)
```

But main strength of this package is in combining created layouts with specific ratio:

```{r combine, message = FALSE, results = 'hide', fig.width = 4, fig.height = 2, tidy = FALSE}
# lay will be 3 times wider that lay2
# all ascects in lay and lay2 will be preserved
cl = layColBind(lay, lay2, widths = c(3, 1))
layShow(cl)
```

You can also create even more complicated layouts:

```{r complicated, message = FALSE, results = 'hide', fig.width = 4, fig.height = 2}
lay3 <- layCreate(matrix(1:2))
lay4 <- layRowBind(cl, lay3, heights = c(5, 2))
layShow(lay4)

lay5 <- layColBind(lay4, lay3, widths = c(5, 2))
layShow(lay5)
```


## Split layout

You can create very complicated layouts by splitting one field:

```{r split, message = FALSE, results = 'hide', fig.width = 4, fig.height = 2, tidy = FALSE, fig.keep='all'}
library(customLayout)
lay <- layCreate(
  matrix(1:4, nc = 2),
  widths = c(3, 2),
  heights = c(2, 1))
layShow(lay)
lay2 <- layCreate(
  matrix(1:4, nc = 2),
  widths = c(3, 5),
  heights = c(2, 4))
layShow(lay2)

# Split field 4 from lay into lay2:
slay <- laySplitField(lay, lay2, field = 4)
layShow(slay)
```


## Example session:

```{r example,message = FALSE, results = 'hide', fig.width = 10, fig.height = 6}
library(customLayout)
par(mar = c(3, 2, 2, 1))
lay  <- layCreate(
  matrix(1:4, nc = 2),
  widths = c(3, 2),
  heights = c(2, 1))
lay2 <- layCreate(matrix(1:3))
cl   <- layColBind(lay, lay2, widths = c(3, 1))

laySet(cl) # initialize drawing area

set.seed(123)
plot(1:100 + rnorm(100))
plot(rnorm(100), type = "l")
hist(rnorm(500))
acf(rnorm(100))
pie(c(3, 4, 6), col = 2:4)
pie(c(3, 2, 7), col = 2:4 + 3)
pie(c(5, 4, 2), col = 2:4 + 6)
```
