customLayout
========================================================

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(cache=TRUE)
```

`customLayout` is a simple extension of basic `layout` function from `R`.

## Instalation:

`customLayout` is available only via GitHub and can be installed using `devtools`:

```{r, eval=FALSE}
require(devtools)
install_github("customLayout",username="zzawadz")
```


## Basic functionality:

You can simply create layouts like in base R:
```{r,message=FALSE,results='hide',fig.width=4,fig.height=2,fig.keep='all',fig.pos='H'}
require(customLayout)
lay = createLayout(matrix(1:4,nc=2),widths=c(3,2),heights=c(2,1))
layoutShow(lay)
lay2 = createLayout(matrix(1:4,nc=2),widths=c(3,5),heights=c(2,4))
layoutShow(lay2)
```
But main strength of this package is in combining created layouts with specific ratio:
```{r,message=FALSE,results='hide',fig.width=4,fig.height=2,tidy=FALSE}
# lay will be 3 times wider that lay2
# all ascects in lay and lay2 will be preserved
cl = colBind(lay,lay2, widths=c(3,1))
layoutShow(cl)
```
You can also create even more complicated layouts:
```{r,message=FALSE,results='hide',fig.width=4,fig.height=2}
lay3 = createLayout(matrix(1:2))
lay4 = rowBind(cl,lay3, heights=c(5,2))
layoutShow(lay4)

lay5 = colBind(lay4,lay3, widths=c(5,2))
layoutShow(lay5)
```

## Example session:

```{r,message=FALSE,results='hide',fig.width=10,fig.height=6}
require(customLayout)
par(mar = c(3,2,2,1))
lay = createLayout(matrix(1:4,nc=2),widths=c(3,2),heights=c(2,1))
lay2 = createLayout(matrix(1:3))
cl = colBind(lay,lay2, widths=c(3,1))
setLayout(cl) # initialize drawing area
plot(1:100+rnorm(100))
plot(rnorm(100), type = "l")
hist(rnorm(500))
acf(rnorm(100))
pie(c(3,4,6),col = 2:4)
pie(c(3,2,7),col = 2:4+3)
pie(c(5,4,2),col = 2:4+6)
```

## Store plots in Layout

You can create `Layout` with plots and then combine it with another `Layout`.

```{r,message=FALSE,results='hide',fig.width=10,fig.height=6,tidy=FALSE}
require(customLayout)
par(mar = c(3,2,2,1))


# listPlots capture code for drawing plots
plots = listPlot(plot(1:100+rnorm(100)),
plot(rnorm(100), type = "l"),
hist(rnorm(500)),
acf(rnorm(100)))
lay = createLayout(matrix(1:4,nc=2),widths=c(3,2),heights=c(2,1),plots=plots)


# renderPlots create drawing area and execute code stored in Layout
renderPlots(lay)

pie_plots = listPlot(pie(c(3,4,6),col = 2:4),
    pie(c(3,2,7),col = 2:4+3),
    {
      # you can capture code in {}
      x = sample(sample(1:10,3))
      pie(x,col = 2:4+6)
      }
)

lay2 = createLayout(matrix(1:3),plots=pie_plots)
renderPlots(lay2)

cl = colBind(lay,lay2, widths=c(3,1))
# render combined layout
renderPlots(cl)
```



